Single Bivalent Level Analysis

### 2020-01-24 doc/Bivalent_Analysis.Rmd
coffee shop work

Bivalent level traits and metrics have been added in the src/Setup_BivData.Rmd script. These observations are from the automated image analysis algorithm and have been curated (threw out incorrect algorithm output).

@@

-

@@

              #sis.co.ten!!
              mean.siscoten = mean(SisCoTen, na.rm=TRUE)
)

@@

#MAKE the same version of the TABLE but for the longest_biv_Dataset (when I get around to making it)

#code if males are high or low?
```

@@

first.SC.plot <- ggplot(Curated_BivData, aes(y = chromosomeLength, x= strain, fill= hand.foci.count))+geom_jitter(aes(alpha=0.2))+geom_boxplot(aes(alpha=0.5))+  ggtitle("SC Length for Bivalents")+facet_wrap(sex~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)
first.SC.plot
#add boxplots for mouse level means
#add second x axis, hand.foci.count

@@

@@

Above is the model we are testing evolution of SC lengths.  remember for the MM use the table **mouse.avs_4MM**

remember the long biv data set in **real.long.bivData**

@@

@@

### Normalized 1CO positions

@@

```{r cent.dist, echo=FALSE}

cent.dist_boxplots <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere distance")+ ylim(c(0, 150))

@@

>>>>

Bivalent level traits and metrics have been added in the src/Setup_BivData.Rmd script. These observations are from the automated image analysis algorithm and have been curated (threw out incorrect algorithm output). The MLH1 data file is also loaded into this file.

@@

I think a formula something like this can be applied to adjust for XX

- rate of bivalent segmentation /* rate of XX, 5% /* mean SC length for 3rd longest bivalent / total SC area (by bivalent) = proportion of SC araea due to XX

@@

              #sis.co.ten!!
              mean.siscoten = mean(SisCoTen, na.rm=TRUE),

              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
)

@@

#MAKE the same version of the TABLE but for the longest_biv_Dataset (when I get around to making it)
#code if males are high or low?
```

@@

first.SC.plot <- ggplot(Curated_BivData, aes(y = chromosomeLength, x= strain, fill= hand.foci.count))+geom_jitter(aes(alpha=0.2))+geom_boxplot(aes(alpha=0.5))+  ggtitle("SC Length for Bivalents")+facet_wrap(sex~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(30,200))
first.SC.plot
#add boxplots for mouse level means
#add second x axis, hand.foci.count

@@

#### Mixed Model Tests

@@

The evolution will be tested with the subsp fixed effect and sexual dimorphism of the trait will be tested with the sex and interaction effects.

remember for the MM use the table **mouse.avs_4MM**  remember the long biv data set in **real.long.bivData**

@@

explain the difference between the raw and PER measures. Why should I use /examine the IFD.PER?

@@

# Normalized 1CO positions

@@

```{r MM.telo, echo=FALSE}

lme_MM_telo.dist <- lme(mean.telo.dist ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
as.data.frame( round(  summary(lme_MM_telo.dist)$tTable, 6 ) )

lmer_MM_telo.dist <- lmer(mean.telo.dist ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
#anova(lmer_MM_siscoten)

rand_telo.dist_MM <- exactRLRT(lmer_MM_telo.dist)
rand_telo.dist_MM
```

```{r cent.dist, echo=FALSE}
cent.dist_boxplots <- ggplot(Curated_BivData, aes(x= strain, y=dis.cent ) )+ geom_boxplot(aes( fill=hand.foci.count, group=strain), alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("raw centromere distance")+ ylim(c(0, 150))

@@

```{r MM.cent.dist, echo=FALSE}
lme_MM_cent.dist <- lme(mean.cent.dist ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )
as.data.frame( round(  summary(lme_MM_cent.dist)$tTable, 6 ) )

lmer_MM_cent.dist <- lmer(mean.cent.dist ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)

exactRLRT(lmer_MM_cent.dist)
```
